---
title: "Foxlights_Modelling"
author: "Lara Volski"
date: "2022-07-11"
output: html_document
---

# Part 1) Setting up

### 1.1 Load Packages
```{r}
library(MASS)
library(lme4)
library(ggplot2)
library(dplyr)
library(car)
library(tidyr)
```

### 1.2 Read in record table CSV
```{r record.table}
record.table.all <- read.csv("record_table_0min_deltaT_2020-06-06.csv")
```


### 1.3 create new column for cameras by taking the folder name and reducing to first three characters
```{r new.columns}
record.table.all$Camera <- strtrim(record.table.all$Camera_folder, 5)
record.table.all$Phase <- substr(record.table.all$Camera_folder, 7, 12)
record.table.all$Foxlight <- substr(record.table.all$Camera_folder, 14, 17)
```


### 1.4 Set the time interval (photos that occured within 15 min will count as one detection)
```{r fifteen.min}
record.table.15min <- record.table.all[record.table.all$delta.time.secs >= 900 |  # because 900 seconds = 15 minutes
                                             record.table.all$delta.time.secs == 0, ]
```

### 1.5 Import additional csv


*metadata.raster is normally distributed. The closer to 0, the less the value deviates from the norm. 
```{r import.csv}
# import camera phase operation dates
cam_operations <- read.csv("phase_operation_dates_2020.csv", header=T)
# reading in metadata
# metadata <- read.csv("MicrohabitatData.csv", header=T)
metadata.raster <- read.csv("Raster_metadata_Lara.csv", header=T) %>% 
  select(-c(Latitude, Longitude))
# scale raster values
metadata.raster <- metadata.raster %>%
  mutate_if(is.numeric, scale)
```

### 1.6 Calculate RAI
```{r RAI}
# calculate RAI for each period
# change camera, phase, and species to factor
record.table.15min$Camera <- as.factor(record.table.15min$Camera)
record.table.15min$Phase <- as.factor(record.table.15min$Phase)
record.table.15min$Species <- droplevels(as.factor(record.table.15min$Species))
# calculate number of observations of each species at each camera in each phase
rai <- record.table.15min %>%
  dplyr::group_by(Species, Camera, Phase, .drop = FALSE) %>%
  dplyr::summarise(Detections = n()) 
rai2 <- record.table.15min %>%
  dplyr::group_by(Species, .drop = FALSE) %>%
  dplyr::summarise(Detections = n())
# merge with record table
rai <- merge(rai, cam_operations)
# calculate RAI
rai$RAI <- rai$Detections / rai$Operation
# remove records where camera was operating for <10 days
for (i in 1:nrow(rai)) {
  if(rai$Operation[i] < 10) {
    rai$Detections[i] <- NA
    rai$RAI[i] <- NA
  } 
}
# join raster metadata
rai <- left_join(rai, metadata.raster)
#### now just calculate RAI at each camera, ACROSS ALL PHASES
cam_operations.acrossphases <- read.csv("camera_operation_dates_2020.csv")
# calculate number of observations of each species at each camera
rai.acrossphases <- record.table.15min %>%
  dplyr::group_by(Species, Camera, .drop = FALSE) %>%
  dplyr::summarise(Detections = n()) 
# merge with record table
rai.acrossphases <- merge(rai.acrossphases, cam_operations.acrossphases)
# calculate RAI
rai.acrossphases$RAI <- rai.acrossphases$Detections / rai.acrossphases$Operation
# join with metadata
# rai.acrossphases <- left_join(rai.acrossphases, metadata.raster, by = "Camera")
```

# Part 2 - Model selection

Make candidate models and evaluate them!

## Null model/hypothesis - only camera as random effect.
AIC = 329.0
```{r}
fit.null <- glmer.nb(Detections ~ offset(log(Operation)) + (1|Camera), data = rai_bobcat_lv)
summary(fit.null)
```

## Foxlight Family of Models
### Foxlight only
AIC = 330.5
```{r}
fit1 <- glmer.nb(Detections ~ Foxlight + offset(log(Operation)) + (1|Camera), data = rai_bobcat_lv)
summary(fit1)
```

### Sheep only
AIC = 330.5
```{r}
fit2 <- glmer.nb(Detections ~ Sheep_Presence + offset(log(Operation)) + (1|Camera), data = rai_bobcat_lv)
summary(fit2)
```

### Phase only 
AIC = 335.8
```{r}
fit3 <- glmer.nb(Detections ~ Phase + offset(log(Operation)) + (1|Camera), data = subset(rai, Species == "LYRU"), control=glmerControl(optimizer="bobyqa"))
summary(fit3)
```

### Foxlight + Sheep
AIC = 334.8
```{r}
fit4 <- glmer.nb(Detections ~ Foxlight + Sheep_Presence + offset(log(Operation)) + (1|Camera), data = subset(rai, Species == "LYRU"))
summary(fit4)
vif(fit4)
```

### Foxlight * Sheep
AIC = 333.6
```{r}
fit5 <- glmer.nb(Detections ~ Foxlight * Sheep_Presence + offset(log(Operation)) + (1|Camera), data = subset(rai, Species == "LYRU"))
summary(fit5)
vif(fit5)
```

### Foxlight + Phase
AIC = 337.7
```{r}
fit6 <- glmer.nb(Detections ~ Foxlight + Phase + offset(log(Operation)) + (1|Camera), data = subset(rai, Species == "LYRU"), control=glmerControl(optimizer="bobyqa"))
summary(fit6)
```

### Foxlight * Phase
AIC = 343.3 
```{r}
fit7 <- glmer.nb(Detections ~ Foxlight * Phase + offset(log(Operation)) + (1|Camera), data = subset(rai, Species == "LYRU"), control=glmerControl(optimizer="bobyqa"))
summary(fit7)
```

### Foxlight + Sheep + Phase
AIC = 339.4
```{r}
fit8 <- glmer.nb(Detections ~ Foxlight + Phase + Sheep_Presence + offset(log(Operation)) + (1|Camera), data = subset(rai, Species == "LYRU"), control=glmerControl(optimizer="bobyqa"))
summary(fit8)
```

## Habitat Variables

### Elevation
AIC = 326.6
```{r}
fit9 <- glmer.nb(Detections ~ Elevation + offset(log(Operation)) + (1|Camera), data = rai_bobcat_lv, control=glmerControl(optimizer="bobyqa"))
summary(fit9)
```

### Ruggedness
AIC = 328.4
```{r}
fit10 <- glmer.nb(Detections ~ Rugged.121 + offset(log(Operation)) + (1|Camera), data = rai_bobcat_lv, control=glmerControl(optimizer="bobyqa"))
summary(fit10)
```

### Foxlight * Elevation
AIC = 330.3
```{r}
fit11 <- glmer.nb(Detections ~ Foxlight * elevation.clean + offset(log(Operation)) + (1|Camera), data = subset(rai, Species == "LYRU"), control=glmerControl(optimizer="bobyqa"))
summary(fit11)
vif(fit11)
```

### Foxlight * Ruggedness121
AIC = 329.9
```{r}
fit12 <- glmer.nb(Detections ~ Foxlight * Rugged.121 + offset(log(Operation)) + (1|Camera), data = rai_bobcat_lv, control=glmerControl(optimizer="bobyqa"))
summary(fit12)
vif(fit12)
```

### Sheep_Presence * Elevation
AIC = 326.9
```{r}
fit13 <- glmer.nb(Detections ~ Sheep_Presence * elevation.clean + offset(log(Operation)) + (1|Camera), data = subset(rai, Species == "LYRU"), control=glmerControl(optimizer="bobyqa"))
summary(fit13)
vif(fit13)
```

### Sheep_Presence * Ruggedness
AIC = 331.4
```{r}
fit14 <- glmer.nb(Detections ~ Sheep_Presence * rugged121.clean + offset(log(Operation)) + (1|Camera), data = subset(rai, Species == "LYRU"), control=glmerControl(optimizer="bobyqa"))
summary(fit14)
vif(fit14)
```

## Dogs
AIC = 328.1
```{r}
rai_bobcat2 <- read.csv("rai_bobcat_lv.csv", header=T)
fit15 <- glmer.nb(Detections ~ I(Ind_Dogs^2) + Ind_Dogs + offset(log(Operation)) + (1|Camera), data = rai_bobcat2)
summary(fit15)
```

## Dogs * Rugged.121
AIC = 326.5
```{r}
fit16 <- glmer.nb(Detections ~ Ind_Dogs * Rugged.121 + offset(log(Operation)) + (1|Camera), data = rai_bobcat2)
summary(fit16)
vif(fit16)
```

## Dogs Scaled
AIC = 328.1
```{r}
fit17 <- glmer.nb(Detections ~ Dogs_Scaled + offset(log(Operation)) + (1|Camera), data = rai_bobcat2)
summary(fit17)
vif(fit17)
```

## Dogs_RAI
AIC = 328.1
```{r}
fit18 <- glmer.nb(Detections ~ Dog_RAI + offset(log(Operation)) + (1|Camera), data = rai_bobcat_lv)
summary(fit18)
vif(fit18)
```

## Road Distance
AIC = 331
```{r}
fit19 <- glmer.nb(Detections ~ road.dist.clean + offset(log(Operation)) + (1|Camera), data = rai_bobcat_lv, control=glmerControl(optimizer="bobyqa"))
summary(fit19)
```
## Dog_RAI * Elevation * Foxlight * Ruggedness
AIC = 331
# convergence error
```{r}
fit20 <- glmer.nb(Detections ~ Dog_RAI * Elevation * Foxlight * Rugged.121 + offset(log(Operation)) + (1|Camera), data = rai_bobcat_lv, control=glmerControl(optimizer="bobyqa"))
summary(fit20)
vif(fit20)
```


# Part 3 - Dredge Models

```{r}
rai_bobcat_lv <- read.csv("rai_bobcat_lv.csv")

## Full lmer model
foxlights.full.model <- lmer(Detections ~ Foxlight + Sheep_Presence + Dog_RAI + Dogs_Scaled + Elevation +
                     rugged9.clean + rugged25.clean + rugged49.clean + rugged81.clean +
                     Rugged.121  + Slope + hq.dist.clean + Phase +
                     road.dist.clean + viewshed.clean +
                        offset(log(Operation)) + (1|Camera),
                      data = rai_bobcat_lv, na.action = "na.fail")

AIC(foxlights.full.model) ## AIC is 437.096
vif(foxlights.full.model) ## All except Foxlight, Sheep_Presence, and Dogs_RAI have VIF > 2

## Reduced lmer model
reduced.model.foxlights <- lmer(Detections ~ Foxlight + Sheep_Presence + Dog_RAI + Elevation +
                     Rugged.121  + Slope + Phase +
                     road.dist.clean + viewshed.clean +
                        offset(log(Operation)) + (1|Camera),
                      data = rai_bobcat_lv, na.action = "na.fail")

AIC(reduced.model.foxlights) ## AIC is 459.8778
vif(reduced.model.foxlights) ## All have a VIF < 2

#Now, to dredge! First install and load the MuMIn package
install.packages("MuMIn")
library(MuMIn)

#Then dredging is as simple as...
dredge.reduced.model.foxlights <- dredge(reduced.model.foxlights)
summary(reduced.model.foxlights)

# to get R2 (both ways work)
# option 1
r.squaredGLMM(reduced.model.foxlights)
# option 2
library(sjPlot)
tab_model(reduced.model.foxlights)


```


