---
title: "Foxlights_Modelling"
author: "Lara Volski"
date: "2022-07-11"
output: html_document
---

# Part 1) Setting up

### 1.1 Load Packages
```{r}
library(MASS)
library(lme4)
library(ggplot2)
library(dplyr)
library(car)
library(tidyr)
```

### 1.2 Read in record table CSV
```{r record.table}
record.table.all <- read.csv("record_table_0min_deltaT_2020-06-06.csv")
```


### 1.3 create new column for cameras by taking the folder name and reducing to first three characters
```{r new.columns}
record.table.all$Camera <- strtrim(record.table.all$Camera_folder, 5)
record.table.all$Phase <- substr(record.table.all$Camera_folder, 7, 12)
record.table.all$Foxlight <- substr(record.table.all$Camera_folder, 14, 17)
```


### 1.4 Set the time interval (photos that occured within 15 min will count as one detection)
```{r fifteen.min}
record.table.15min <- record.table.all[record.table.all$delta.time.secs >= 900 |  # because 900 seconds = 15 minutes
                                             record.table.all$delta.time.secs == 0, ]
```

### 1.5 Import additional csv


*metadata.raster is normally distributed. The closer to 0, the less the value deviates from the norm. 
```{r import.csv}
# import camera phase operation dates
cam_operations <- read.csv("phase_operation_dates_2020.csv", header=T)
# reading in metadata
# metadata <- read.csv("MicrohabitatData.csv", header=T)
metadata.raster <- read.csv("Raster_metadata_Lara.csv", header=T) %>% 
  select(-c(Latitude, Longitude))
# scale raster values
metadata.raster <- metadata.raster %>%
  mutate_if(is.numeric, scale)
```

### 1.6 Calculate RAI
```{r RAI}
# calculate RAI for each period
# change camera, phase, and species to factor
record.table.15min$Camera <- as.factor(record.table.15min$Camera)
record.table.15min$Phase <- as.factor(record.table.15min$Phase)
record.table.15min$Species <- droplevels(as.factor(record.table.15min$Species))
# calculate number of observations of each species at each camera in each phase
rai <- record.table.15min %>%
  dplyr::group_by(Species, Camera, Phase, .drop = FALSE) %>%
  dplyr::summarise(Detections = n()) 
rai2 <- record.table.15min %>%
  dplyr::group_by(Species, .drop = FALSE) %>%
  dplyr::summarise(Detections = n())
# merge with record table
rai <- merge(rai, cam_operations)
# calculate RAI
rai$RAI <- rai$Detections / rai$Operation
# remove records where camera was operating for <10 days
for (i in 1:nrow(rai)) {
  if(rai$Operation[i] < 10) {
    rai$Detections[i] <- NA
    rai$RAI[i] <- NA
  } 
}
# join raster metadata
rai <- left_join(rai, metadata.raster)
#### now just calculate RAI at each camera, ACROSS ALL PHASES
cam_operations.acrossphases <- read.csv("camera_operation_dates_2020.csv")
# calculate number of observations of each species at each camera
rai.acrossphases <- record.table.15min %>%
  dplyr::group_by(Species, Camera, .drop = FALSE) %>%
  dplyr::summarise(Detections = n()) 
# merge with record table
rai.acrossphases <- merge(rai.acrossphases, cam_operations.acrossphases)
# calculate RAI
rai.acrossphases$RAI <- rai.acrossphases$Detections / rai.acrossphases$Operation
# join with metadata
# rai.acrossphases <- left_join(rai.acrossphases, metadata.raster, by = "Camera")
```

# Part 2 - Model selection

Make candidate models and evaluate them!

## Null model/hypothesis - only camera as random effect.
AIC = 329.0
```{r}
fit.null <- glmer.nb(Detections ~ offset(log(Operation)) + (1|Camera), data = subset(rai, Species == "LYRU"))
summary(fit.null)
```

## Foxlight Family of Models
### Foxlight only
AIC = 330.5
```{r}
fit1 <- glmer.nb(Detections ~ Foxlight + offset(log(Operation)) + (1|Camera), data = subset(rai, Species == "LYRU"))
summary(fit1)
```

### Sheep only
AIC = 330.5
```{r}
fit2 <- glmer.nb(Detections ~ Sheep_Presence + offset(log(Operation)) + (1|Camera), data = subset(rai, Species == "LYRU"))
summary(fit2)
```

### Phase only 
AIC = 335.8
```{r}
fit3 <- glmer.nb(Detections ~ Phase + offset(log(Operation)) + (1|Camera), data = subset(rai, Species == "LYRU"), control=glmerControl(optimizer="bobyqa"))
summary(fit3)
```

### Foxlight + Sheep
AIC = 334.8
```{r}
fit4 <- glmer.nb(Detections ~ Foxlight + Sheep_Presence + offset(log(Operation)) + (1|Camera), data = subset(rai, Species == "LYRU"))
summary(fit4)
vif(fit4)
```

### Foxlight * Sheep
AIC = 333.6
```{r}
fit5 <- glmer.nb(Detections ~ Foxlight * Sheep_Presence + offset(log(Operation)) + (1|Camera), data = subset(rai, Species == "LYRU"))
summary(fit5)
vif(fit5)
```

### Foxlight + Phase
AIC = 337.7
```{r}
fit6 <- glmer.nb(Detections ~ Foxlight + Phase + offset(log(Operation)) + (1|Camera), data = subset(rai, Species == "LYRU"), control=glmerControl(optimizer="bobyqa"))
summary(fit6)
```

### Foxlight * Phase
AIC = 343.3 
```{r}
fit7 <- glmer.nb(Detections ~ Foxlight * Phase + offset(log(Operation)) + (1|Camera), data = subset(rai, Species == "LYRU"), control=glmerControl(optimizer="bobyqa"))
summary(fit7)
```

### Foxlight + Sheep + Phase
AIC = 339.4
```{r}
fit8 <- glmer.nb(Detections ~ Foxlight + Phase + Sheep_Presence + offset(log(Operation)) + (1|Camera), data = subset(rai, Species == "LYRU"), control=glmerControl(optimizer="bobyqa"))
summary(fit8)
```

## Habitat Variables

### Distance to Water
AIC = 323.1
```{r}
fit9 <- glmer.nb(Detections ~ water.dist.clean + offset(log(Operation)) + (1|Camera), data = subset(rai, Species == "LYRU"), control=glmerControl(optimizer="bobyqa"))
summary(fit9)
```

### Elevation
AIC = 326.6
```{r}
fit10 <- glmer.nb(Detections ~ elevation.clean + offset(log(Operation)) + (1|Camera), data = subset(rai, Species == "LYRU"), control=glmerControl(optimizer="bobyqa"))
summary(fit10)
```

### Ruggedness
AIC = 328.4
```{r}
fit11 <- glmer.nb(Detections ~ rugged121.clean + offset(log(Operation)) + (1|Camera), data = subset(rai, Species == "LYRU"), control=glmerControl(optimizer="bobyqa"))
summary(fit11)
```

### Foxlight * Distance to Water
AIC = 326
```{r}
fit12 <- glmer.nb(Detections ~ Foxlight * water.dist.clean + offset(log(Operation)) + (1|Camera), data = subset(rai, Species == "LYRU"), control=glmerControl(optimizer="bobyqa"))
summary(fit12)
vif(fit12)
```

### Foxlight * Elevation
AIC = 330.3
```{r}
fit13 <- glmer.nb(Detections ~ Foxlight * elevation.clean + offset(log(Operation)) + (1|Camera), data = subset(rai, Species == "LYRU"), control=glmerControl(optimizer="bobyqa"))
summary(fit13)
vif(fit13)
```

### Foxlight * Ruggedness121
AIC = 329.9
```{r}
fit14 <- glmer.nb(Detections ~ Foxlight * rugged121.clean + offset(log(Operation)) + (1|Camera), data = subset(rai, Species == "LYRU"), control=glmerControl(optimizer="bobyqa"))
summary(fit14)
vif(fit14)
```

### Sheep_Presence * Distance to Water
AIC = 325.7
```{r}
fit15 <- glmer.nb(Detections ~ Sheep_Presence * water.dist.clean + offset(log(Operation)) + (1|Camera), data = subset(rai, Species == "LYRU"), control=glmerControl(optimizer="bobyqa"))
summary(fit15)
vif(fit15)
```

### Sheep_Presence * Elevation
AIC = 326.9
```{r}
fit16 <- glmer.nb(Detections ~ Sheep_Presence * elevation.clean + offset(log(Operation)) + (1|Camera), data = subset(rai, Species == "LYRU"), control=glmerControl(optimizer="bobyqa"))
summary(fit16)
vif(fit16)
```

### Sheep_Presence * Ruggedness
AIC = 331.4
```{r}
fit17 <- glmer.nb(Detections ~ Sheep_Presence * rugged121.clean + offset(log(Operation)) + (1|Camera), data = subset(rai, Species == "LYRU"), control=glmerControl(optimizer="bobyqa"))
summary(fit17)
vif(fit17)
```

## Dogs
AIC = 328.1
```{r}
rai_bobcat2 <- read.csv("rai_bobcat_lv.csv", header=T)
fit18 <- glmer.nb(Detections ~ I(Ind_Dogs^2) + Ind_Dogs + offset(log(Operation)) + (1|Camera), data = rai_bobcat2)
summary(fit18)
```

## Dogs + Dist.Water
AIC = 320.8
```{r}
fit19 <- glmer.nb(Detections ~ Ind_Dogs + Dist.Water + offset(log(Operation)) + (1|Camera), data = rai_bobcat2)
summary(fit19)
vif(fit19)
```

## Dogs * Dist.Water
AIC = 322.6
VIF is over 4!!!
```{r}
fit20 <- glmer.nb(Detections ~ Ind_Dogs * Dist.Water + offset(log(Operation)) + (1|Camera), data = rai_bobcat2)
summary(fit20)
vif(fit20)
```

## Dogs * Rugged.121
AIC = 326.5
```{r}
fit21 <- glmer.nb(Detections ~ Ind_Dogs * Rugged.121 + offset(log(Operation)) + (1|Camera), data = rai_bobcat2)
summary(fit21)
vif(fit21)
```

## Dogs Scaled
AIC = 328.1
```{r}
fit22 <- glmer.nb(Detections ~ Dogs_Scaled + offset(log(Operation)) + (1|Camera), data = rai_bobcat2)
summary(fit22)
vif(fit22)
```

## Dogs Scaled * Dist_Water
AIC = 322.6
VIF for Dogs_Scaled = 18.7768
```{r}
fit23 <- glmer.nb(Detections ~ Dogs_Scaled * Dist.Water + offset(log(Operation)) + (1|Camera), data = rai_bobcat2)
summary(fit23)
vif(fit23)
```

## Dogs Scaled + Dist_Water
AIC = 320.8
VIF < 4
```{r}
fit24 <- glmer.nb(Detections ~ Dogs_Scaled + Dist.Water + offset(log(Operation)) + (1|Camera), data = rai_bobcat2)
summary(fit24)
vif(fit24)
```

## 
AIC = 320.8
VIF < 4
```{r}
fit25 <- glmer.nb(Detections ~ Rugged.121 + offset(log(Operation)) + (1|Camera), data = rai_bobcat2, control=glmerControl(optimizer="bobyqa"))
summary(fit25)
vif(fit25)
```


# Part 3 - Dredge Models

```{r}
rai_bobcat2 <- read.csv("rai_bobcat_lv.csv")
## Full lmer model
full.model <- lmer(Detections ~ Foxlight + Sheep_Presence + Dogs_Scaled + Elevation +
                     rugged9.clean + rugged25.clean + rugged49.clean + rugged81.clean +
                     Rugged.121 + Dist.Water + Slope + Fence.Dist + hq.dist.clean +
                     road.dist.clean + viewshed.clean + ndvi.16.clean.1 +
                     veg.edges.dist.clean + chap.edges.dist.clean + 
                        offset(log(Operation)) + (1|Camera),
                      data = rai_bobcat2, na.action = "na.fail")
AIC(full.model) ## AIC is 427.4306
vif(full.model) ## All except Foxlight, Sheep_Presence, and Dogs_Scaled have VIF > 4
full.model.reduced <- lmer(Detections ~ Foxlight + Sheep_Presence + Dogs_Scaled +
                               Dist.Water + Slope + Fence.Dist + road.dist.clean +
                               veg.edges.dist.clean + chap.edges.dist.clean + 
                               (1|Camera), 
                               data = rai_bobcat2, na.action = "na.fail")
AIC(full.model.reduced) ## AIC is 454.5723
vif(full.model.reduced) ## All except Sheep_Presence have a VIF < 2, Sheep_Presence is 2.023585
#Now, to dredge! First install and load the MuMIn package
#Then dredging is as simple as...
Dredge.Table <- dredge(full.model.reduced)
trial.model <- lmer(Detections ~ Foxlight + Sheep_Presence + Dogs_Scaled +
                      Dist.Water + Slope + Fence.Dist +
                     road.dist.clean + viewshed.clean +
                        offset(log(Operation)) + (1|Camera),
                      data = rai_bobcat2, na.action = "na.fail")
AIC(trial.model) ## AIC is 427.4306
vif(trial.model)
## last trial
trial.model <- lmer(Detections ~ Foxlight + Dogs_Scaled +
                      Dist.Water + Slope + Fence.Dist +
                     road.dist.clean + viewshed.clean +
                        offset(log(Operation)) + (1|Camera),
                      data = rai_bobcat2, na.action = "na.fail")
AIC(trial.model) ## AIC is 454.5723
vif(trial.model) ## All except Sheep_Presence have a VIF < 2, Sheep_Presence is 2.023585
Dredge.Table2 <- dredge(trial.model)
