---
title: "LGD_Models_UsingNB"
author: "Lara Volski"
date: "2022-11-26"
output: html_document
---

Read in Data + load packages
```{r}
subset.rai <- read.csv("subset.rai.csv")
library(sjPlot)
```

Null Model
AIC = 2575.127
```{r}
## Null model/hypothesis - only camera as random effect.
fit.null <- glmer.nb(Count ~ offset(log(Operation)) + (1|Camera), data = subset.rai)
summary(fit.null)
tab_model(fit.null)
confint(fit.null, level=0.95)
```

Sheep Presence (Binary, Indicative of dog)
AIC = 2577.1 
```{r}
fit.sheep <- glmer.nb(Count ~ Sheep + offset(log(Operation)) + (1|Camera), data = subset.rai)
summary(fit.sheep)
tab_model(fit.sheep)
confint(fit.sheep, level=0.95)
```

Dog Presence (Binary)
AIC = 2576.6, Estimate = 0.19
```{r}
fit.dog <- glmer.nb(Count ~ Dog + offset(log(Operation)) + (1|Camera), data = subset.rai, control=glmerControl(optimizer="bobyqa"))
summary(fit.dog)
tab_model(fit.dog)
```

Dog Presence (RAI)
AIC = 2576.8, Estimate = 0.662034 # Failure to converge!
```{r}
# Scaling Dog_RAI

mean(subset.rai$Dog_RAI) #0.006537058
sd(subset.rai$Dog_RAI) # 0.05177796

(0 - 0.006537058) / 0.05177796 # -0.1262517

fit.dog.rai <- glmer.nb(Count ~ Dog_RAI_Scaled + offset(log(Operation)) + (1|Camera), data = subset.rai, control=glmerControl(optimizer="bobyqa"))
summary(fit.dog.rai)
tab_model(fit.dog.rai)

ggplot(subset.rai, aes(x = Bobcat_RAI, y = Dog_RAI_Scaled)) +
  geom_smooth()
```

Individual Dog Count (Scaled)
AIC = 2576.9, Estimate = 0.022456 # Failure to converge!
```{r}
fit.dog.scaled <- glmer.nb(Count ~ Dogs_Scaled + offset(log(Operation)) + (1|Camera), data = subset.rai, control=glmerControl(optimizer="bobyqa"))
summary(fit.dog.scaled)

```

Fire
AIC = 2577.1, Estimate = -0.128
```{r}
fit.fire <- glmer.nb(Count ~ Fire + offset(log(Operation)) + (1|Camera), data = subset.rai)
summary(fit.fire)
```

Season
AIC = 2532.8, Estimate = negative summer, positive winter
```{r}
fit.season <- glmer.nb(Count ~ Season + offset(log(Operation)) + (1|Camera), data = subset.rai)
summary(fit.season)
tab_model(fit.season)
```

Ruggedness.9
AIC = 2577.1, Estimate = -0.00342
```{r}
fit.rugged9 <- glmer.nb(Count ~ rugged9.clean + offset(log(Operation)) + (1|Camera), data = subset.rai)
summary(fit.rugged9)
```

Ruggedness.121
AIC = 2568.2, Estimate = 0.5970
```{r}
fit.rugged121 <- glmer.nb(Count ~ rugged121.clean + offset(log(Operation)) + (1|Camera), data = subset.rai)
summary(fit.rugged121)
tab_model(fit.rugged121)
```

Distance from road
AIC = 2575.6, Estimate = 0.2697
```{r}
fit.road <- glmer.nb(Count ~ road.dist.clean + offset(log(Operation)) + (1|Camera), data = subset.rai)
summary(fit.road)
```

Elevation
AIC = 2577.0, Estimate = 0.08634
```{r}
fit.elevation <- glmer.nb(Count ~ elevation.clean + offset(log(Operation)) + (1|Camera), data = subset.rai)
summary(fit.elevation)
```

Slope
AIC = 2577.0, Estimate = 0.08295
```{r}
fit.slope <- glmer.nb(Count ~ slope.clean + offset(log(Operation)) + (1|Camera), data = subset.rai)
summary(fit.slope)
```

Vegetation Type
AIC = 2578.1, Estimate = grassland and woodland are both negative
```{r}
fit.veg <- glmer.nb(Count ~ vegetation.coarser.clean + offset(log(Operation)) + (1|Camera), data = subset.rai)
summary(fit.veg)
```

# Combining Estimates
Season + Ruggedness
AIC = 2525.8, Marginal/Conditional R^2 = 0.072 / 0.206, CI = 
```{r}
fit.1 <- glmer.nb(Count ~ Season + rugged121.clean + offset(log(Operation)) + (1|Camera), data = subset.rai)
summary(fit.1)
vif(fit.1)
tab_model(fit.1)
fit.1.CIs <- confint(fit.1, level=0.95)
```

Season + Ruggedness + Dog_RAI
AIC = 2527.785, Marginal/Conditional R^2 = 0.072 / 0.206, CI = 
```{r}
fit.2 <- glmer.nb(Count ~ Season + rugged121.clean + Dog_RAI_Scaled + offset(log(Operation)) + (1|Camera), data = subset.rai)
vif(fit.2)
summary(fit.2)
tab_model(fit.2)
fit.2.CIs <- confint(fit.2, level=0.95)
```

Season * Dog_RAI
AIC = 2539.206, Marginal/Conditional R^2 = 0.031 / 0.209
```{r}
fit.3 <- glmer.nb(Count ~ Dog_RAI * Season + offset(log(Operation)) + (1|Camera), data = subset.rai)
vif(fit.3) # Over 2?
summary(fit.3)
tab_model(fit.3)
```

Ruggedness * Dog_RAI
AIC = 2571.856, Marginal/Conditional R^2 = 0.045 / 0.183
```{r}
fit.4 <- glmer.nb(Count ~ rugged121.clean * Dog_RAI + offset(log(Operation)) + (1|Camera), data = subset.rai)
vif(fit.4) # Over 2?
summary(fit.4)
tab_model(fit.4)
```

Ruggedness * Dog_RAI
AIC = 2577.247, Marginal/Conditional R^2 = 0.045 / 0.183
```{r}
fit.5 <- glmer.nb(Count ~ Sheep * Fire + offset(log(Operation)) + (1|Camera), data = subset.fire.rai)
vif(fit.5) # Over 2?
tab_model(fit.5)
summary(fit.4)
tab_model(fit.4)
```


